Recursive Solutions to \\First-Order Model Counting

### 2023-01-30 doc/paper/paper.tex
A new short talk (and some other things)

% See https://www.overleaf.com/learn/latex/theorems_and_proofs
% for a nice explanation of how to define new theorems, but keep
% in mind that the amsthm package is already included in this
% template and that you must *not* alter the styling.
% \newtheorem{example}{Example}
% \newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}

@@

% TODO: double check how the university names should be formatted and the emails
% TODO: read the formatting instructions and their LaTeX source
% TODO: make sure there are no overflow errors
% TODO: for arbitrary instances of '2022' (e.g., in citations), make it '2023'
\author{%
Paulius Dilkas$^1$\and
Vaishak Belle$^2$\\
\affiliations
$^1$National University of Singapore, Singapore\\
$^2$University of Edinburgh, UK\\
\emails
paulius.dilkas@nus.edu.sg,
vbelle@ed.ac.uk
}

@@

\newcommand{\crefrangeconjunction}{--}
\newcommand\pfun{\mathrel{\ooalign{\hfil$\mapstochar\mkern5mu$\hfil\cr$\to$\cr}}}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\newcommand{\FOtwo}{$\mathsf{FO}^{2}$}
\newcommand{\FOthree}{$\mathsf{FO}^{3}$}
\newcommand{\SFO}{$\mathsf{S}^{2}\mathsf{FO}^{2}$}
\newcommand{\SRU}{$\mathsf{S}^{2}\mathsf{RU}$}
\newcommand{\Uone}{$\mathsf{U}_{1}$}
\newcommand{\Ctwo}{$\mathsf{C}^{2}$}
\newcommand{\IFO}{$\mathsf{I}\mathsf{FO}^{2}$}
\DeclareMathOperator{\CR}{\textsc{CR}}
\DeclareMathOperator{\GDR}{\textsc{GDR}}
\DeclareMathOperator{\Reff}{\textsc{Ref}}
\DeclareMathOperator{\dom}{dom}
\DeclareMathOperator{\Doms}{Doms}
\DeclareMathOperator{\Vars}{Vars}
\SetKwProg{Fn}{Function}{:}{}
\SetKwFunction{identifyRecursion}{r}
\SetKwFunction{generateMaps}{genMaps}
\makeatletter
\newcommand{\nosemic}{\renewcommand{\@endalgocfline}{\relax}}% Drop semi-colon ;
\newcommand{\dosemic}{\renewcommand{\@endalgocfline}{\algocf@endline}}% Reinstate semi-colon ;
\newcommand{\pushline}{\Indp}% Indent
\newcommand{\popline}{\Indm\dosemic}% Undent
\makeatother
\Crefname{algocf}{Algorithm}{Algorithms}
\crefname{line}{line}{lines}

@@

% TODO: fix doom emacs warnings
Traditionally in computational complexity theory, a problem is \emph{tractable}
if it can be solved in time polynomial in the instance size. The equivalent
notion in (W)FOMC is liftability. A (W)FOMC instance is \emph{(domain-)liftable}
if it can be solved in time polynomial in the size(s) of the
domain(s)~\cite{jaeger2012liftability}. Over more than a decade, many classes of
instances were shown to be
liftable~\cite{DBLP:conf/kr/BremenK21,DBLP:conf/nips/KazemiKBP16,DBLP:conf/lics/KuusistoL18,DBLP:journals/jair/Kuzelka21}.
First, Van den Broeck~\shortcite{DBLP:conf/nips/Broeck11} showed that the class
of all sentences of FOL with up to two variables (denoted \FOtwo{}) is liftable.
Then Beame et al.~\shortcite{DBLP:conf/pods/BeameBGS15} proved that there exists
a sentence with three variables for which FOMC is $\#\P_{1}$-complete (i.e.,
\FOthree{} is not liftable). Since these two results came out, most of the
research on (W)FOMC focused on developing faster solutions for the \FOtwo{}
fragment~\cite{DBLP:conf/uai/BremenK21,DBLP:conf/aaai/MalhotraS22} and defining
new liftable fragments. These fragments include \SFO{} and
\SRU{}~\cite{DBLP:conf/nips/KazemiKBP16},
\Uone{}~\cite{DBLP:conf/lics/KuusistoL18}, \FOtwo{} with tree
axioms~\cite{DBLP:conf/kr/BremenK21}, and \Ctwo{} (i.e., the two-variable
fragment with counting
quantifiers)~\cite{DBLP:journals/jair/Kuzelka21,DBLP:conf/aaai/MalhotraS22}. On
the empirical front, there are several open-source implementations of exact
WFOMC algorithms: \textsc{ForcLift}~\cite{DBLP:conf/ijcai/BroeckTMDR11},
probabilistic theorem proving~\cite{DBLP:journals/cacm/GogateD16}, and
\textsc{L2C}~\cite{DBLP:conf/kr/KazemiP16}. Approximate counting is supported by
\textsc{ApproxWFOMC}~\cite{DBLP:conf/ijcai/BremenK20} as well as
\textsc{ForcLift}~\cite{DBLP:conf/uai/BroeckCD12} and probabilistic theorem
proving~\cite{DBLP:journals/cacm/GogateD16}.

@@

However, none of the publicly available exact (W)FOMC algorithms can efficiently
compute functions as simple as a factorial.\footnote{There is previous
  work~\cite{DBLP:journals/jair/Kuzelka21,DBLP:conf/ijcai/Broeck16} showing how
  FOMC can compute factorials in principle, but an algorithm that can lift such
  an instance in practice does not exist.} We claim that this shortcoming is due
to the inability of these algorithms to construct recursive solutions. The topic
of recursion in the context of WFOMC has been studied before but in limited
ways. Barv{\'{\i}}nek et al.~\shortcite{DBLP:conf/ilp/BarvinekB0ZK21} use WFOMC
to generate numerical data that is then used to conjecture recurrence relations
that explain that data. Van den Broeck~\shortcite{DBLP:conf/nips/Broeck11}
introduced the idea of \emph{domain recursion}. Intuitively, domain recursion
partitions a domain of size $n$ into a single explicitly named constant and the
remaining domain of size $n-1$. However, many stringent conditions are enforced
to ensure that the search for a tractable solution always terminates.

@@

TODO: connecting sentence?

\begin{proposition}[Correctness of GDR]
  Let $\phi$ be the formula used as input to \cref{alg:domainrecursion} and
  $\phi'$ be one of the formulas constructed by the algorithm. Then
  $\phi \equiv \phi'$.
\end{proposition}
\begin{proof}
  TODO
  Assume that \cref{line:condition} selects domain $\Omega \in \mathcal{D}$.
  \[
    \forall X \in \Omega\text{. } \psi \equiv \psi[x/\{\, X \,\}] \land \forall X \in \Omega\text{. } X \ne x \implies \psi
  \]
\end{proof}

@@

@@

@@

% TODO: in the acknowledgements, mention that most of the work was done at UoE
% TODO: should the appendix be in the same file or not?

>>>>

% See https://www.overleaf.com/learn/latex/theorems_and_proofs
% for a nice explanation of how to define new theorems, but keep
% in mind that the amsthm package is already included in this
% template and that you must *not* alter the styling.
% \newtheorem{example}{Example}
% \newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{proposition}{Proposition}
\newtheorem{claim}{Claim}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}

@@

% TODO: double check how the university names should be formatted and the emails
% TODO: read the formatting instructions and their LaTeX source
% TODO: make sure there are no overflow errors
% TODO: for arbitrary instances of '2022' (e.g., in citations), make it '2023'
% TODO: maybe mention the term 'data complexity'
% TODO: fix doom emacs warnings
\author{%
Paulius Dilkas$^1$\and
Vaishak Belle$^2$\\
\affiliations
$^1$National University of Singapore, Singapore\\
$^2$University of Edinburgh, UK\\
\emails
paulius.dilkas@nus.edu.sg,
vbelle@ed.ac.uk
}

@@

\newcommand{\crefrangeconjunction}{--}
\newcommand\pfun{\mathrel{\ooalign{\hfil$\mapstochar\mkern5mu$\hfil\cr$\to$\cr}}}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\newcommand{\FOtwo}{$\mathsf{FO}^{2}$}
\newcommand{\FOthree}{$\mathsf{FO}^{3}$}
\newcommand{\SFO}{$\mathsf{S}^{2}\mathsf{FO}^{2}$}
\newcommand{\SRU}{$\mathsf{S}^{2}\mathsf{RU}$}
\newcommand{\Uone}{$\mathsf{U}_{1}$}
\newcommand{\Ctwo}{$\mathsf{C}^{2}$}
\newcommand{\IFO}{$\mathsf{I}\mathsf{FO}^{2}$}
\DeclareMathOperator{\CR}{\textsc{CR}}
\DeclareMathOperator{\GDR}{\textsc{GDR}}
\DeclareMathOperator{\Reff}{\textsc{Ref}}
\DeclareMathOperator{\dom}{dom}
\DeclareMathOperator{\Doms}{Doms}
\DeclareMathOperator{\MC}{MC}
\DeclareMathOperator{\Vars}{Vars}
\SetKwProg{Fn}{Function}{:}{}
\SetKwFunction{identifyRecursion}{r}
\SetKwFunction{generateMaps}{genMaps}
\makeatletter
\newcommand{\nosemic}{\renewcommand{\@endalgocfline}{\relax}}% Drop semi-colon ;
\newcommand{\dosemic}{\renewcommand{\@endalgocfline}{\algocf@endline}}% Reinstate semi-colon ;
\newcommand{\pushline}{\Indp}% Indent
\newcommand{\popline}{\Indm\dosemic}% Undent
\makeatother
\Crefname{algocf}{Algorithm}{Algorithms}
\crefname{line}{line}{lines}

@@

Traditionally in computational complexity theory, a problem is \emph{tractable}
if it can be solved in time polynomial in the instance size. The equivalent
notion in (W)FOMC is liftability. A (W)FOMC instance is \emph{(domain-)liftable}
if it can be solved in time polynomial in the size(s) of the
domain(s)~\cite{jaeger2012liftability}. Over more than a decade, many classes of
instances were shown to be
liftable~\cite{DBLP:conf/kr/BremenK21,DBLP:conf/nips/KazemiKBP16,DBLP:conf/lics/KuusistoL18,DBLP:journals/jair/Kuzelka21}.
First, Van den Broeck~\shortcite{DBLP:conf/nips/Broeck11} showed that the class
of all sentences of FOL with up to two variables (denoted \FOtwo{}) is liftable.
Then Beame et al.~\shortcite{DBLP:conf/pods/BeameBGS15} proved that there exists
a sentence with three variables for which FOMC is $\#\P_{1}$-complete (i.e.,
\FOthree{} is not liftable). Since these two results came out, most of the
research on (W)FOMC focused on developing faster solutions for the \FOtwo{}
fragment~\cite{DBLP:conf/uai/BremenK21,DBLP:conf/aaai/MalhotraS22} and defining
new liftable fragments. These fragments include \SFO{} and
\SRU{}~\cite{DBLP:conf/nips/KazemiKBP16},
\Uone{}~\cite{DBLP:conf/lics/KuusistoL18}, \FOtwo{} with tree
axioms~\cite{DBLP:conf/kr/BremenK21}, and \Ctwo{} (i.e., the two-variable
fragment with counting
quantifiers)~\cite{DBLP:journals/jair/Kuzelka21,DBLP:conf/aaai/MalhotraS22}. On
the empirical front, there are several open-source implementations of exact
WFOMC algorithms: \textsc{ForcLift}~\cite{DBLP:conf/ijcai/BroeckTMDR11},
probabilistic theorem proving~\cite{DBLP:journals/cacm/GogateD16}, and
\textsc{L2C}~\cite{DBLP:conf/kr/KazemiP16}. Approximate counting is supported by
\textsc{ApproxWFOMC}~\cite{DBLP:conf/ijcai/BremenK20} as well as
\textsc{ForcLift}~\cite{DBLP:conf/uai/BroeckCD12} and probabilistic theorem
proving~\cite{DBLP:journals/cacm/GogateD16}.
% TODO: add FastWFOMC (only for \FOtwo{})

@@

However, none of the publicly available exact (W)FOMC algorithms can efficiently
compute functions as simple as a factorial.\footnote{There is some previous work
  showing how FOMC can compute
  factorials~\cite{DBLP:journals/jair/Kuzelka21,DBLP:conf/kr/BremenK21,DBLP:conf/ijcai/Broeck16}.
  One of them~\cite{DBLP:conf/ijcai/Broeck16} is an example `by hand' that does
  not directly correspond to any algorithm.
  Another~\cite{DBLP:journals/jair/Kuzelka21} proves that the problem is
  liftable but contains no experimental results. The
  third~\cite{DBLP:conf/kr/BremenK21} contains experimental results performed
  with an algorithm, which is not publicly available.} We claim that this
shortcoming is due to the inability of these algorithms to construct recursive
solutions. The topic of recursion in the context of WFOMC has been studied
before but in limited ways. Barv{\'{\i}}nek et
al.~\shortcite{DBLP:conf/ilp/BarvinekB0ZK21} use WFOMC to generate numerical
data that is then used to conjecture recurrence relations that explain that
data. Van den Broeck~\shortcite{DBLP:conf/nips/Broeck11} introduced the idea of
\emph{domain recursion}. Intuitively, domain recursion partitions a domain of
size $n$ into a single explicitly named constant and the remaining domain of
size $n-1$. However, many stringent conditions are enforced to ensure that the
search for a tractable solution always terminates.

@@

\begin{proposition}[Correctness of $\GDR$]
  Let $\phi$ be the formula used as input to \cref{alg:domainrecursion} and
  $\phi'$ be one of the formulas constructed by the algorithm. Then
  $\phi \equiv \phi'$.
\end{proposition}
\begin{proof} % TODO: finish (Proof technique: if one is true then the other one will be true and vice versa)
  Let $\Omega \in \mathcal{D}$ be the domain selected by
  \cref{line:condition} of the algorithm.
  \begin{claim}\label{claim:logic}
  \[
    \forall X \in \Omega\text{. } \psi \equiv \psi[x/\{\, X \,\}] \land \forall X \in \Omega\text{. } X \ne x \implies \psi
  \]
  \end{claim}
  \begin{claim}
    Iteration of \cref{claim:logic} for all variables in some set of variables
    $V$ produces the same clauses as subset enumeration on
    \cref{line:conditions}.
  \end{claim}
\end{proof}

@@

\begin{proposition}[Correctness of $\CR$]
\end{proposition}
\begin{proof} % TODO: finish (Proof technique: if one is true then the other one will be true and vice versa)
  \[
    \forall X \in \Omega \ni x\text{. } X \ne x \implies \psi \equiv \forall X \in \Omega \setminus \{\, x \,\}\text{. } \psi
  \]
\end{proof}

@@

% TODO: maybe something other than MC? Introduce the notation and the role of \sigma.
% TODO: add an argument about why this recursion is (or can easily be made) finite
\begin{proposition}[Correctness of $\Reff$]
\end{proposition}
\begin{proof} % TODO: finish (Proof technique: show a bijection between the sets of models)
  For a given formula $\phi$, \cref{alg:trycache} attempts to find a formula
  $\psi$ and a map $\rho\colon \Doms(\psi) \to \Doms(\phi)$ such that
  $\MC(\psi, \sigma \circ \rho\colon \Doms(\psi) \to \mathbb{N}_{0}) = \MC(\phi, \sigma\colon \Doms(\phi) \to \mathbb{N}_{0})$.
\end{proof}

@@

% TODO: in the acknowledgements, mention that most of the work was done at UoE
% TODO: should the appendix be in the same file or not?
% TODO: also mention that most of the work for this paper was done at UoE

### 2023-01-30 doc/talks/3_long/talk.tex
A new short talk (and some other things)

\usetikzlibrary{shapes}
\usetikzlibrary{arrows}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{positioning}
\usetikzlibrary{overlay-beamer-styles}
\usetikzlibrary{shapes.callouts}

@@

% 15 min introduction
% 25 min presentation
% 5 min summary
% 20-25 slides
% right now it's 15 slides and 20 min
% TODO: use \alert instead of \structure

@@

% TODO: must attribute vecteezy
\begin{frame}{What Computers Can and Cannot Do}
  \begin{tikzpicture}[remember picture,overlay]
    \node[right=0.7cm] at (current page.west) (stickman) {\includegraphics[height=0.5\textheight]{stickman.png}};
    \node<-4>[left,shift=({0, -2})] at (current page.east) (computer) {\includegraphics[height=0.5\textheight]{happy_computer}};
    \node<2>[ellipse callout, draw, right=0cm, callout relative pointer={(-1, -0.5)}] at (stickman.north east) {$\int_{-\infty}^{\infty} e^{-t^{2}}\,\mathrm{d}t$};
    \node<2>[ellipse callout, draw, left=-0.5cm, below=0.5cm] at (computer.north west) {$\sqrt{\pi}$};
    \node<3>[ellipse callout, draw, right=-1cm, callout relative pointer={(-2.5, -0.5)}] at (stickman.north east) {Produce a schedule for the nurses at NUH.};
    \node<3>[ellipse callout, draw, below=0.5cm, shift=({-0.5,0}), callout relative pointer={(0.5, -0.25)}] at (computer.north west) {\includegraphics[width=0.1\textwidth]{schedule}};
    \node<4>[ellipse callout, draw, right=0cm, callout relative pointer={(-1, -0.15)}] at (stickman.north east) {Describe this picture. \includegraphics[width=0.25\textwidth]{cat.jpg}};
    \node<4>[ellipse callout, draw, below=0.5cm, shift=({-0.5,0.5}), callout relative pointer={(0.5, -0.5)}] at (computer.north west) {It's a cat with its eyes closed.};
    \node<5>[ellipse callout, draw, right=0cm, callout relative pointer={(-1, -0.15)}, text width=5cm] at (stickman.north east) {If I shuffle a deck of \alert{$n$} cards, how many possible outcomes are there?};
    \node<5>[left,shift=({-1, -1})] at (current page.east) (computer2) {\includegraphics[height=0.5\textheight]{overheating_computer}};
  \end{tikzpicture}
  % 1. compute the integral
  % 2. job scheduling
  % 3. picture description
  % 4. count matchings
\end{frame}

\begin{frame}{Flavours of Counting}
\end{frame}

\begin{frame}{Some Elementary Counting}
  \begin{exampleblock}{A Counting Problem}
    Suppose this room has \structure{$n$} seats, and there are
    \structure{$m \le n$} people in the audience. How many ways are there to
    seat everyone?
  \end{exampleblock}

@@

  \pause
  \alert{Answer:} \structure{$n^{\underline{m}} = n \cdot (n-1)\cdots(n-m+1)$}.

  Note: this problem is equivalent to counting \structure{$[m] \to [n]$} injections.

@@

\begin{frame}{Let's Express This Problem in Logic!}
  \begin{itemize}
  \item Let \structure{$M$} and \structure{$N$} be sets (i.e., \alert{domains}) such that \structure{$|M| = m$}, and \structure{$|N| = n$}
  \item Let \structure{$P \subseteq M \times N$} be a relation (i.e., \alert{predicate}) over sets \structure{$M$} and \structure{$N$}
  \item We can describe all of the constraints in first-order logic:
    \begin{itemize}
    \item \pause each attendee gets a seat (i.e., at least one seat)
      \[
      \forall x \in M\text{. }\exists y \in N\text{. }P(x, y)
      \]
    \item \pause one person cannot occupy multiple seats
      \[
      \forall x \in M\text{. }\forall y, z \in N\text{. }P(x, y) \land P(x, z) \Rightarrow y=z
      \]
    \item \pause one seat cannot accommodate multiple attendees
      \[
      \forall w, x \in M\text{. }\forall y \in N\text{. }P(w, y) \land P(x, y) \Rightarrow w=x
      \]
    \end{itemize}
  \end{itemize}
  \pause
  The first two sentences constrain \structure{$P$} to be a function, and the last one makes it injective.
\end{frame}
% one could also add probabilities to these sentences to say that, e.g., there is a 1% probability of somebody taking up two seats, somebody sitting on somebody else's lap, somebody sitting on the floor, etc.

\begin{frame}{Overview of the Problem}
  \begin{itemize}
    \item \alert{First-order model counting} (FOMC) is the problem of counting
          the models of a sentence in first-order logic.
    \item The \alert{(symmetric) weighted} variation of the problem adds weights
          (e.g., probabilities) to predicates.
          \begin{itemize}
            \item It is used for efficient \alert{probabilistic inference} in
                  relational models such as Markov logic networks.
          \end{itemize}
    \item None of the (implemented) (W)FOMC algorithms are able to count, e.g.,
          \alert{injective} and \alert{bijective} functions.
  \end{itemize}
  \begin{block}{Claim}
    This shortcoming can be addressed via support for (almost arbitrary)
    \alert{recursive functions}.
  \end{block}
\end{frame}

\section{Main Content (TODO: remove later)}

@@

\section{Summary (TODO: remove later)}

>>>>

\usetikzlibrary{shapes}
\usetikzlibrary{arrows}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{positioning}
\usetikzlibrary{overlay-beamer-styles}
\usetikzlibrary{shapes.callouts}
\usetikzlibrary{cd}

@@

\newcommand{\WMC}{\textup{WMC}}
\newcommand{\WFOMC}{\textup{(W)FOMC}}
\newcommand{\WMI}{\textup{WMI}}
\newcommand{\WFOMI}{\textup{WFOMI}}
\newcommand{\AMC}{\textup{AMC}}
\newcommand{\pbp}{\textup{PBP}}
\newcommand{\SPPP}{\textup{SP}}

% 15 min introduction
% 20-25 min presentation
% 5-10 min summary
% 20-25 slides
% before it was 15 slides and 20 min
% TODO: use \alert instead of \structure
% TODO: replace capital latin letters with capital Greek letters

@@

\begin{frame}{What Computers Can and Cannot Do}
  \begin{tikzpicture}[remember picture,overlay]
    \node[right=0.7cm] at (current page.west) (stickman) {\includegraphics[height=0.5\textheight]{stickman.png}};
    \node<-4>[left,shift=({0, -2})] at (current page.east) (computer) {\includegraphics[height=0.5\textheight]{happy_computer}};
    \node<2>[ellipse callout, draw, right=0cm, callout relative pointer={(-1, -0.5)}] at (stickman.north east) {$\int_{-\infty}^{\infty} e^{-t^{2}}\,\mathrm{d}t$};
    \node<2>[ellipse callout, draw, left=-0.5cm, below=0.5cm] at (computer.north west) {$\sqrt{\pi}$};
    \node<3>[ellipse callout, draw, right=-1cm, callout relative pointer={(-2.5, -0.5)}] at (stickman.north east) {Produce a schedule for the nurses at NUH.};
    \node<3>[ellipse callout, draw, below=0.5cm, shift=({-0.5,0}), callout relative pointer={(0.5, -0.25)}] at (computer.north west) {\includegraphics[width=0.1\textwidth]{schedule}};
    \node<4>[ellipse callout, draw, right=0cm, callout relative pointer={(-1, -0.15)}] at (stickman.north east) {Describe this picture. \includegraphics[width=0.25\textwidth]{cat.jpg}};
    \node<4>[ellipse callout, draw, below=0.5cm, shift=({-0.5,0.5}), callout relative pointer={(0.5, -0.5)}] at (computer.north west) {It's a cat with its eyes closed.};
    \node<5>[ellipse callout, draw, right=0cm, callout relative pointer={(-1, -0.15)}, text width=5cm] at (stickman.north east) {If I shuffle a deck of \alert{$n$} cards, how many possible outcomes are there?};
    \node<5>[left,shift=({-1, -1})] at (current page.east) (computer2) {\includegraphics[height=0.5\textheight]{overheating_computer}};
    \node[shift=({1.3, 0.1})] at (current page.south west) {\tiny \color{lightgray}{Vector art by Vecteezy.com}};
  \end{tikzpicture}
\end{frame}
% TODO: replace cat with 'Generate a picture of a monkey with a funny hat'

\begin{frame}[fragile]{Many Ways to Count}
  \begin{center}
    \begin{tikzcd}
      \WMI{} & \WFOMI{} \\
      \#\SAT{}/\WMC{} & \WFOMC{} \\
      \AMC{}, \pbp{} & \SPPP{}
    \end{tikzcd}
  \end{center}
\end{frame}
% maybe mention #CSP and other non-SAT ways of counting

\begin{frame}{Some Elementary Counting}
  \begin{exampleblock}{A Counting Problem}
    Suppose this room has \alert{$n$} seats, and there are \alert{$m \le n$}
    people in the audience. How many ways are there to seat everyone?
  \end{exampleblock}

@@

  \pause
  \alert{Answer:} $n^{\underline{m}} = n \cdot (n-1)\cdots(n-m+1)$.

  Note: this problem is equivalent to counting \alert{$[m] \to [n]$} injections.

@@

\begin{frame}{Let's Express This Problem in Logic!}
  \begin{itemize}
    \item Let \alert{$M$} and \alert{$N$} be sets (i.e., domains)
    \begin{itemize}
      \item such that \alert{$|M| = m$}, and \alert{$|N| = n$}.
    \end{itemize}
    \item Let \alert{$P \subseteq M \times N$} be a relation (i.e., predicate)
          over \alert{$M$} and \alert{$N$}.
    \item We can describe all of the constraints in first-order logic:
    \begin{itemize}
      \item \pause each attendee gets a seat (i.e., at least one seat)
      \[
        \forall x \in M\text{. }\exists y \in N\text{. }P(x, y)
      \]
      \item \pause one person cannot occupy multiple seats
      \[
        \forall x \in M\text{. }\forall y, z \in N\text{. }P(x, y) \land P(x, z) \Rightarrow y=z
      \]
      \item \pause one seat cannot accommodate multiple attendees
      \[
        \forall w, x \in M\text{. }\forall y \in N\text{. }P(w, y) \land P(x, y) \Rightarrow w=x
      \]
    \end{itemize}
  \end{itemize}
  \pause The first two sentences constrain \alert{$P$} to be a function, and the
  last one makes it injective.
\end{frame}
% one could also add probabilities to these sentences to say that, e.g., there is a 1% probability of somebody taking up two seats, somebody sitting on somebody else's lap, somebody sitting on the floor, etc.

\begin{frame}{Overview of the Problem}
  \begin{itemize}
    \item \alert{First-order model counting} (FOMC) is the problem of counting
          the models of a sentence in first-order logic.
    \item The \alert{(symmetric) weighted} variation of the problem adds weights
          (e.g., probabilities) to predicates.
          \begin{itemize}
            \item It is used for efficient \alert{probabilistic inference} in
                  relational models such as Markov logic networks.
          \end{itemize}
    \item None of the (implemented) (W)FOMC algorithms are able to count, e.g.,
          \alert{injective} and \alert{bijective} functions.
  \end{itemize}
  \begin{block}{Claim}
    This shortcoming can be addressed via support for \alert{recursive functions}.
  \end{block}
\end{frame}

% TODO: maybe meat makes more sense than sandwich? Or a picture of recursion, e.g., fractals?
{ % all template changes are local to this group.
    \setbeamertemplate{navigation symbols}{}
    \begin{frame}<article:0>[plain]
        \begin{tikzpicture}[remember picture,overlay]
            \node[at=(current page.center)] {
                \includegraphics[keepaspectratio, width=\paperwidth]{sandwich.jpg}
            };
        \end{tikzpicture}
     \end{frame}
}

@@

{ % all template changes are local to this group.
    \setbeamertemplate{navigation symbols}{}
    \begin{frame}<article:0>[plain]
        \begin{tikzpicture}[remember picture,overlay]
            \node[at=(current page.center)] {
                \includegraphics[keepaspectratio, height=\paperheight]{zoom.jpg}
            };
        \end{tikzpicture}
     \end{frame}
}
